Performance benefits of Intel\textsuperscript{\textregistered}~Optane\texttrademark~DC persistent memory for the parallel processing of large neuroimaging data

### 2019-12-17 paper.tex
Final fixes

Whereas Operating Systems and popular Big Data engines, such as MapReduce~\cite{mapreduce} and Apache Spark~\cite{spark}, have
incorporated software solutions to limit data transfers (e.g. writeback cache, in-memory computing,
data locality, and lazy evaluation), hardware has also adapted to the growing datasets.
One such improvement is the concept of placing persistent storage directly on the Dual
In-line Memory Module (DIMM), thereby reducing the latency of accessing data on 
flash storage devices. While the latency of these devices is improved, the bandwidth
remains the same. However, as noted by \cite{nvdimms}, a severe performance degradation
can be experienced by having memory traffic and I/O placed on the same bus.

@@

For the 20~$\mu$m \bigbrain, we used the same application, however in this case, it was executed
using both Apache Spark and GNU Parallel. The configuration was also generally the same, having 
experiments using both 25 and 96 processes. Repeating the experiments using a larger dataset would enable
us to determine the effects when Optane DCPMM would have to be partially relied upon due to insufficient DRAM
space. Moreover, with such large datasets, it is likely that Big Data frameworks would be used rather than
more traditional parallelism frameworks. Using Spark, we can evaluate how Big Data Frameworks perform on
different storage devices.

@@

As can be seen in Figure~\ref{fig:makespan-25cpus}, DRAM (App Direct mode) is the 
most efficient, with a makespan of 7.6s. Optane in Memory mode is a close second
with a makespan of approximately 10s. As caching was disabled in App Direct mode, 
Optane DCPMM in this mode is approximately 
12x slower than Optane in Memory mode. Isilon in both Memory mode and App Direct 
mode is more than 120x slower than Optane DCPMM in Memory mode. Other than for Optane DCPMM, 
there is no significant difference between Memory mode and App Direct executions.

@@

For App Direct mode applications (although also applicable to Memory mode), the 
choice of using the writeback cache depends on total amount of available memory, application
memory usage and I/O. Should the application be able to use the page cache without ever being
I/O-throttled, using a write-back cache will significantly improve performance. This can 
be seen in Figure~\ref{fig:stacked-25cpus}, for instance, where the local SSD is writing
entirely to memory, significantly reducing the total write time. However, should write operations
become throttled, this may slow down the write task duration. In such cases,
it may be preferable to set them up as write-through devices. As Big Data neuroimaging applications
tend to generate temporary data files that are larger than input size, it is likely that maintaining the
devices as write-through is preferable. Furthermore, despite the fact that Optane DCPMM in App Direct mode does
not leverage the writeback cache like the local SSD, is it still found to be superior in performance.
Therefore, while the performance of Optane DCPMM may not be enhanced by DRAM, it is still superior to that of
an SSD using the writeback cache and scalable network storage.

@@

Storage performance when scaling the application from 25 to 96 cores could have been improved using a
NUMA-aware application, since memory is tied to the CPUs. Should a task requiring memory
stored on one CPU socket be sent to another CPU, this would incur some overheads, reducing overall application
and storage performance. Furthermore, Intel optimized Python and NumPy libraries could have been used to improve 
performance.

>>>>

Whereas Operating Systems and popular Big Data engines, such as MapReduce~\cite{mapreduce} and Apache Spark~\cite{spark}, have
incorporated software solutions to limit data transfers (e.g. writeback cache, in-memory computing,
data locality, and lazy evaluation), hardware has also adapted to the growing datasets.
One such improvement is the concept of placing persistent storage directly on the Dual
In-line Memory Module (DIMM), thereby reducing the latency of accessing data on 
storage devices. While the latency of these devices is improved, the bandwidth
remains the same. However, as noted by \cite{nvdimms}, a severe performance degradation
can be experienced by having memory traffic and I/O placed on the same bus.

@@

For the 20~$\mu$m \bigbrain, we used the same application, however in this case, it was executed
using both Apache Spark and GNU Parallel. The configuration was also generally the same, having 
experiments using both 25 and 96 processes. Repeating the experiments with the 20~$\mu$m BigBrain would enable
us to determine the effects when Optane DCPMM would have to be partially relied upon due to insufficient DRAM
space. Moreover, with such large datasets, it is likely that Big Data frameworks would be used rather than
more traditional parallelism frameworks. Using Spark, we can evaluate how Big Data Frameworks perform on
different storage devices.

@@

As can be seen in Figure~\ref{fig:makespan-25cpus}, DRAM (App Direct mode) is the 
most efficient, with a makespan of 7.6s (all makespan values are available \href{https://github.com/big-data-lab-team/paper-memory-storage/blob/master/experiments/bigbrain-incrementation/results/makespan.csv}{here}).
Optane DCPMM in Memory mode is a close second
with a makespan of 10.2s. As caching was disabled in App Direct mode, 
Optane DCPMM in this mode is approximately 
12x slower than Optane DCPMM in Memory mode. Isilon in both Memory mode and App Direct 
mode is more than 120x slower than Optane DCPMM in Memory mode. Other than for Optane DCPMM, 
there is no significant difference between Memory mode and App Direct executions.

@@

For App Direct mode applications (although also applicable to Memory mode), the 
choice of using the writeback cache depends on total amount of available memory, application
memory usage and I/O. Should the application be able to use the page cache without ever being
I/O-throttled, using a write-back cache will significantly improve performance. This can 
be seen in Figure~\ref{fig:stacked-25cpus}, for instance, where the local SSD is writing
entirely to memory, significantly reducing the total write time. However, should write operations
become throttled, this may slow down the write task duration. In such cases,
it may be preferable to set them up as write-through devices. In any case,
 despite the fact that Optane DCPMM in App Direct mode does
not leverage the writeback cache like the local SSD, is it still found to be superior in performance.
Therefore, while the performance of Optane DCPMM may not be enhanced by DRAM, it is still superior to that of
an SSD using the writeback cache and scalable network storage.

@@

Storage performance when scaling the application from 25 to 96 CPU threads could have been improved using a
NUMA-aware application, since memory is tied to the CPUs. Should a task requiring memory
stored on one CPU socket be sent to another CPU, this would incur some overheads, reducing overall application
and storage performance. Furthermore, Intel optimized Python and NumPy libraries could have been used to improve 
performance.
